---
title: "Mada_climate_analysis"
author: "Ke_Fang"
date: "2023-10-30"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r warning=FALSE, message=F}
library(tidyverse)
library(lme4)
library(ordinal)
library(readxl)
library(lmerTest)
library(aod)
```

```{r}
df <- read_xlsx("data/data63_forKay.xlsx")
#glimpse(df)
```

```{r}
df <- df %>% 
  mutate(POLICYcc2 = as.numeric(POLICYcc),
         BELIEFcc2 = as.numeric(BELIEFcc),
         SHAREcc2 = as.numeric(SHAREcc),
         SHAREcc2 = as.factor(SHAREcc2),
         WEPTcc2 = as.numeric(WEPTcc),
         WEPTcc2 = factor(WEPTcc2, order = T),
         condName2 = as.factor(condName),
         condName2 = relevel(condName2, ref = "Control"))
#glimpse(df)
```

## Policy support

### Need for random effect

First, we can compare three models: the model only have condition as fixed effect, a model with condition as fixed effect and country as random effect, and a model with both condition and country as fixed effect (econmetric fixed effect model). Three models can be compared using AIC and BIC.

Result:

- For the fixed effect of conditions, all fixed effect's significance remained unchanged expect for BindingMoral. BindingMoral became not significant in the mixed-effect model and econmetric fixed effect model.
- From AIC, the best model is the econmetric fixed effect model
- From BIC, the best model is the mixed-effect model.

```{r}
M <- glm(POLICYcc2 ~ condName2, data = df %>% filter(!is.na(POLICYcc2), !is.na(condName2)), family = gaussian(link = "identity"))
#print(summary(M))
AIC(M)
BIC(M)
```

```{r}
M <- lmer(POLICYcc2 ~ condName2 + (1 | Country), data = df %>% filter(!is.na(POLICYcc2), !is.na(condName2)), REML = F) 
# use ML estimator consistent as glm()
#print(summary(M))
AIC(M)
BIC(M)
```

```{r}
M <- glm(POLICYcc2 ~ condName2 + factor(Country), data = df %>% filter(!is.na(POLICYcc2), !is.na(condName2)), family = gaussian(link = "identity"))
#print(summary(M))
AIC(M)
BIC(M)
```

### Significance for condition variable

If we decide to go with the mixed-effect model (according to BIC standard):

```{r}
M <- lmer(POLICYcc2 ~ condName2 + (1 | Country), data = df %>% filter(!is.na(POLICYcc2), !is.na(condName2)), REML = T)
print(summary(M))
AIC(M)
BIC(M)
```

#### Method 1: Wald Test

Alternatively, a Wald test tests for whether the coefficient of one or a set of coefficient are equal to 0 (null H0: the expectation of the (joint) distribution of beta parameter(s) of interest is/are 0).

**Result: the wald test is significant, suggesting the condition variables as a block has at least one or a combination of them is significantly different from zero.**

```{r}
condName2_indices <- grep("condName2", names(fixef(M)))
wald.test(b = fixef(M),
              	Sigma = vcov(M),
              	Terms = condName2_indices)
```

#### Method 2: LRT test

LRT compare the fit of two nested models. The LRT assesses whether the more complex model provides a significantly better fit to the data than the simpler model, justifying the inclusion of additional parameters.

To perform the LRT test, we fit a alternative model without condition names:

**Result: the significant LRT show that condition is needed for the model. AIC and BIC also support this.**

For fixed effect: 
```{r}
M_alternative <- lmer(POLICYcc2 ~ 1 + (1 | Country), data = df %>% filter(!is.na(POLICYcc2), !is.na(condName2)))
#print(summary(M_alternative))
AIC(M_alternative)
BIC(M_alternative)

anova(M_alternative, M, refit = T) # use refit = T when fixed effect is different
```

## Belief

### Need for random effect

First, we can compare three models: the model only have condition as fixed effect, a model with condition as fixed effect and country as random effect, and a model with both condition and country as fixed effect (econmetric fixed effect model). Three models can be compared using AIC and BIC.

Result:

- For the fixed effect of conditions, all fixed effect's significance remained unchanged expect for PluralIgnorance. PluralIgnorance became  significant in the mixed-effect model and econmetric fixed effect model.
- From AIC, the best model is the econometric fixed effect model
- From BIC, the best model is the mixed-effect model.

```{r}
M <- glm(BELIEFcc2 ~ condName2, data = df %>% filter(!is.na(BELIEFcc2), !is.na(condName2)), family = gaussian(link = "identity"))
#print(summary(M))
AIC(M)
BIC(M)
```

```{r}
M <- lmer(BELIEFcc2 ~ condName2 + (1 | Country), data = df %>% filter(!is.na(BELIEFcc2), !is.na(condName2)), REML = F) 
# use ML estimator consistent as glm()
#print(summary(M))
AIC(M)
BIC(M)
```

```{r}
M <- glm(BELIEFcc2 ~ condName2 + factor(Country), data = df %>% filter(!is.na(BELIEFcc2), !is.na(condName2)), family = gaussian(link = "identity"))
#print(summary(M))
AIC(M)
BIC(M)
```

### Significance for condition variable

If we decide to go with the mixed-effect model (according to BIC standard):

```{r}
M <- lmer(BELIEFcc2 ~ condName2 + (1 | Country), data = df %>% filter(!is.na(BELIEFcc2), !is.na(condName2)), REML = T)
print(summary(M))
AIC(M)
BIC(M)
```

#### Method 1: Wald Test

Alternatively, a Wald test tests for whether the coefficient of one or a set of coefficient are equal to 0 (null H0: the expectation of the (joint) distribution of beta parameter(s) of interest is/are 0).

**Result: the wald test is significant, suggesting the condition variables as a block has at least one or a combination of them is significantly different from zero.**

```{r}
condName2_indices <- grep("condName2", names(fixef(M)))
wald.test(b = fixef(M),
              	Sigma = vcov(M),
              	Terms = condName2_indices)
```

#### Method 2: LRT test

LRT compare the fit of two nested models. The LRT assesses whether the more complex model provides a significantly better fit to the data than the simpler model, justifying the inclusion of additional parameters.

To perform the LRT test, we fit a alternative model without condition names:

**Result: the significant LRT show that condition is needed for the model. AIC and BIC also support this.**

For fixed effect: 
```{r}
M_alternative <- lmer(BELIEFcc2 ~ 1 + (1 | Country), data = df %>% filter(!is.na(BELIEFcc2), !is.na(condName2)))
#print(summary(M_alternative))
AIC(M_alternative)
BIC(M_alternative)

anova(M_alternative, M, refit = T) # use refit = T when fixed effect is different
```


## Sharing

### Need for random effect

First, we can compare three models: the model only have condition as fixed effect, a model with condition as fixed effect and country as random effect, and a model with both condition and country as fixed effect (econmetric fixed effect model). Three models can be compared using AIC and BIC.

Result:

- For the fixed effect of conditions, all fixed effect's significance remained unchanged.
- From AIC, the best model is the econometric fixed effect model.
- From BIC, the best model is the mixed-effect model.

```{r}
M <- glm(SHAREcc2 ~ condName2, data = df %>% filter(!is.na(SHAREcc2), !is.na(condName2)), family = "binomial")
#print(summary(M))
AIC(M)
BIC(M)
```

```{r}
M <- glmer(SHAREcc2 ~ condName2 + (1 | Country), data = df %>% filter(!is.na(SHAREcc2), !is.na(condName2)), family = binomial) 
# use ML estimator consistent as glm()
#print(summary(M))
AIC(M)
BIC(M)
```



```{r}
M <- glm(SHAREcc2 ~ condName2 + factor(Country), data = df %>% filter(!is.na(SHAREcc2), !is.na(condName2)), family = "binomial")
#print(summary(M))
AIC(M)
BIC(M)
```

### Significance for condition variable

If we decide to go with the mixed-effect model (according to BIC standard):

```{r}
M <- glmer(SHAREcc2 ~ condName2 + (1 | Country), data = df %>% filter(!is.na(SHAREcc2), !is.na(condName2)), family = binomial) 
print(summary(M))
AIC(M)
BIC(M)
```

#### Method 1: Wald Test

Alternatively, a Wald test tests for whether the coefficient of one or a set of coefficient are equal to 0 (null H0: the expectation of the (joint) distribution of beta parameter(s) of interest is/are 0).

**Result: the wald test is significant, suggesting the condition variables as a block has at least one or a combination of them is significantly different from zero.**

```{r}
condName2_indices <- grep("condName2", names(fixef(M)))
wald.test(b = fixef(M),
              	Sigma = vcov(M),
              	Terms = condName2_indices)
```

#### Method 2: LRT test

LRT compare the fit of two nested models. The LRT assesses whether the more complex model provides a significantly better fit to the data than the simpler model, justifying the inclusion of additional parameters.

To perform the LRT test, we fit a alternative model without condition names:

**Result: the significant LRT show that condition is needed for the model. AIC and BIC also support this.**

For fixed effect: 
```{r}
M_alternative <- glmer(SHAREcc2 ~ 1 + (1 | Country), data = df %>% filter(!is.na(SHAREcc2), !is.na(condName2)), family = binomial) 
#print(summary(M_alternative))
AIC(M_alternative)
BIC(M_alternative)

anova(M_alternative, M, refit = T) # use refit = T when fixed effect is different
```


## WEPT

### Need for random effect

First, we can compare three models: the model only have condition as fixed effect, a model with condition as fixed effect and country as random effect, and a model with both condition and country as fixed effect (econmetric fixed effect model). Three models can be compared using AIC and BIC.

Result:

- For the fixed effect of conditions, all fixed effect's significance remained unchanged.
- From AIC, the best model is the econometric fixed effect model.
- From BIC, the best model is the mixed-effect model.

```{r}
M <- clm(WEPTcc2 ~ condName2, data = df %>% filter(!is.na(WEPTcc2), !is.na(condName2)), threshold = "equidistant")
#print(summary(M))
AIC(M)
BIC(M)
```

```{r}
M <- clmm(WEPTcc2 ~ condName2 + (1 | Country), data = df %>% filter(!is.na(WEPTcc2), !is.na(condName2)), threshold = "equidistant") 
# use ML estimator consistent as glm()
print(summary(M))
AIC(M)
BIC(M)
```



```{r}
M <- clm(WEPTcc2 ~ condName2 + factor(Country), data = df %>% filter(!is.na(WEPTcc2), !is.na(condName2)), threshold = "equidistant")
#print(summary(M))
AIC(M)
BIC(M)
```

### Significance for condition variable

If we decide to go with the mixed-effect model (according to BIC standard):

```{r}
M <- clmm(WEPTcc2 ~ condName2 + (1 | Country), data = df %>% filter(!is.na(WEPTcc2), !is.na(condName2)), threshold = "equidistant")
print(summary(M))
AIC(M)
BIC(M)
```

#### Method 1: Wald test

The Wald block test is less supported for clmm(), so let's do it by hand

   \[
   W = (\hat{\boldsymbol{\beta}} - \boldsymbol{\beta}_0)^\top \left[\text{Cov}(\hat{\boldsymbol{\beta}})\right]^{-1} (\hat{\boldsymbol{\beta}} - \boldsymbol{\beta}_0)
   \]

   where \( \hat{\boldsymbol{\beta}} \) is the vector of estimated coefficients, \( \boldsymbol{\beta}_0 \) is the vector of coefficients under the null hypothesis (often a vector of zeros), and \( \text{Cov}(\hat{\boldsymbol{\beta}}) \) is the covariance matrix of the estimated coefficients.

The test statistic \( W \) follows, approximately, a chi-square distribution with degrees of freedom equal to the number of constraints being tested (the number of fixed effects being simultaneously evaluated).

**Result: the Wald test is significant p < .001**

```{r}
W = t(M$coefficients[3:13]) %*% solve(vcov(M)[3:13,3:13]) %*% M$coefficients[3:13]

1 - pchisq(W, df = 11)
```

#### Method 2: LRT test

LRT compare the fit of two nested models. The LRT assesses whether the more complex model provides a significantly better fit to the data than the simpler model, justifying the inclusion of additional parameters.

To perform the LRT test, we fit a alternative model without condition names:

**Result: the significant LRT show that condition is needed for the model. AIC and BIC also support this.**

For fixed effect: 
```{r}
M_alternative <- clmm(WEPTcc2 ~ 1 + (1 | Country), data = df %>% filter(!is.na(WEPTcc2), !is.na(condName2)), threshold = "equidistant") 
#print(summary(M_alternative))
AIC(M_alternative)
BIC(M_alternative)

anova(M_alternative, M) # use refit = T when fixed effect is different
```





